---
---
Title: "Comparative Morphometric & Physiological Analysis of *Labeo rohita* and *Catla catla* across West Bengal Market Sites"
Author: "ANANYA SAMADDER | UID No: 22BSZ70039 | HONOURS IN ZOOLOGY"
Date: "`r format(Sys.Date(), '%B %Y')`"

```{r setup, include=FALSE}
# This chunk hides all code and warnings from the final PDF
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE,
  fig.align = "center",
  dpi = 150
)
```

---

# Abstract

This study investigates the morphometric and physiological variation in two commercially important carp species — *Labeo rohita* (Rohu) and *Catla catla* (Catla) — sampled from three market sites in West Bengal: Kalyani, Barasat, and Gayeshpur. Using Fulton's Condition Factor (K), Length–Weight Relationship (LWR) regression, Principal Component Analysis (PCA), Pearson correlation, and ANOVA with Tukey HSD post-hoc tests, we identify a clear **pollution signature** in the Barasat population characterised by stunted growth and reduced condition factors. Our findings confirm that anthropogenic environmental stressors significantly influence fish biology in this region.

---

# Introduction

Major carps such as *L. rohita* and *C. catla* are ecologically and economically vital to West Bengal's aquaculture and fisheries sector. Morphometric variation across spatially separated populations can reveal the impact of localised environmental degradation. Market sites serve as proxies for catchment-area habitat quality, since fish are typically sourced from nearby water bodies.

**Research Objectives:**

1. Quantify physiological health (K) across three market sites.
2. Determine growth patterns (allometry) using Length–Weight Relationships.
3. Detect morphometric divergence using PCA.
4. Validate findings using ANOVA, Tukey HSD, and Pearson correlation.

---

# Materials & Methods

## Data Collection

Fish specimens were collected from three market sites:

| Site | Location | n (Catla) | n (Rohu) |
|------|----------|-----------|----------|
| Kalyani | Nadia District | ~12 | ~12 |
| Barasat | North 24 Parganas | ~10 | ~12 |
| Gayeshpur | Nadia District | ~12 | ~12 |

Measurements recorded: Total Length (TL, cm), Head Length (HL, cm), Body Depth (BD, cm), Dorsal Fin Length (DFL), Pectoral Fin Length (PFL), Anal Fin Length (AFL), and Weight (g).

## Statistical Methods

| Method | Purpose |
|--------|---------|
| Fulton's K | Physiological health index |
| LWR (log–log regression) | Growth allometry (slope b) |
| Pearson Correlation | Bivariate relationships |
| One-Way ANOVA | Site effect on K and Weight |
| Tukey HSD | Post-hoc pairwise comparisons |
| Two-Way ANOVA | Site × Species interaction |
| PCA | Multivariate morphometric divergence |

---

# Setup & Data Loading

```{r libraries}
library(tidyverse)
library(ggpubr)
library(car)
library(multcomp)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(rstatix)
library(knitr)

SITE_COLORS    <- c(Kalyani = "#2196F3", Barasat = "#FF9800", Gayeshpur = "#4CAF50")
SPECIES_COLORS <- c(Rohu = "#1a237e", Catla = "#b71c1c")
SITES          <- c("Kalyani", "Barasat", "Gayeshpur")

theme_dissertation <- function() {
  theme_classic(base_size = 12) +
    theme(
      plot.title    = element_text(face = "bold", size = 13, hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5, colour = "grey40"),
      axis.title    = element_text(face = "bold"),
      legend.title  = element_text(face = "bold"),
      strip.background = element_rect(fill = "grey92", colour = NA),
      strip.text    = element_text(face = "bold")
    )
}
``` 

```{r load-data}
# 1. Load with 'show_col_types = FALSE' to keep the console clean
catla_raw <- read_csv("final_catla.csv", show_col_types = FALSE)
rohu_raw  <- read_csv("final_rohu.csv",  show_col_types = FALSE)

# 2. Combine and FORCE everything to be numeric
# This handles cases where a "space" makes R think a number is "text"
df <- bind_rows(
  catla_raw %>% mutate(Species = "Catla"),
  rohu_raw  %>% mutate(Species = "Rohu")
) %>%
  rename(
    Total_Length        = TL,
    Head_Length         = HL,
    Body_Depth          = BD,
    Dorsal_Fin_Length   = DFL,
    Pectoral_Fin_Length = PFL,
    Anal_Fin_Length     = AFL
  ) %>%
  mutate(
    # Convert to numeric just in case there's a hidden space/char
    Weight       = as.numeric(Weight),
    Total_Length = as.numeric(Total_Length),
    # Morphometric Ratios
    BD_TL  = Body_Depth          / Total_Length,
    HL_TL  = Head_Length         / Total_Length,
    DFL_TL = Dorsal_Fin_Length   / Total_Length,
    PFL_TL = Pectoral_Fin_Length / Total_Length,
    AFL_TL = Anal_Fin_Length     / Total_Length,
    # Condition Factor
    K       = (Weight / Total_Length^3) * 100,
    Site    = factor(Site,    levels = SITES),
    Species = factor(Species, levels = c("Rohu", "Catla"))
  )

# VERIFICATION: If this doesn't say 70, the issue is the file reading step.
cat(sprintf("FINAL DATASET COUNT: %d fish specimens loaded.\n", nrow(df)))
```

## Descriptive Statistics

```{r descriptive_statistics}
library(kableExtra)
df %>%
  group_by(Species, Site) %>%
  summarise(
    n           = n(),
    TL_mean     = round(mean(Total_Length, na.rm = TRUE), 2),
    TL_sd       = round(sd(Total_Length,   na.rm = TRUE), 2),
    Weight_mean = round(mean(Weight, na.rm = TRUE), 2),
    Weight_sd   = round(sd(Weight,   na.rm = TRUE), 2),
    .groups     = "drop"
  ) %>%
  kbl(caption = "Table 1. Descriptive statistics by Species and Site") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#455A64") %>% # Professional Slate Grey header
  column_spec(1, bold = T, border_right = T) %>% # Bolds Species and adds a separator
  column_spec(2, italic = T) # Italics the Site names for a professional look
```

---

# Results

## Fulton's Condition Factor (K)

Fulton's K was calculated as:

$$K = \frac{W}{L^3} \times 100$$

where W = weight (g) and L = total length (cm).

```{r condition-factor}

# 1. Calculate K
df <- df %>%
  mutate(K = (Weight / Total_Length^3) * 100)

# 2. Generate Styled Table
df %>%
  group_by(Species, Site) %>%
  summarise(
    K_mean = round(mean(K, na.rm=TRUE), 4),
    K_sd   = round(sd(K,   na.rm=TRUE), 4),
    .groups = "drop"
  ) %>%
  kbl(caption = "Table 2. Mean Condition Factor (K) by Species and Site") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#607D8B") %>% # Professional Blue-Grey header
  column_spec(3, bold = T, color = "darkblue") # Highlights the Mean K values for quick reading
```

```{r fig-condition-factor, fig.cap="Figure 1. Fulton's Condition Factor across sites. Dashed line = overall mean K."}
overall_K <- mean(df$K, na.rm = TRUE)

ggplot(df, aes(x = Site, y = K, fill = Site)) +
  geom_boxplot(outlier.shape = 21, alpha = 0.75, width = 0.6, colour = "grey30") +
  geom_hline(yintercept = overall_K, linetype = "dashed",
             colour = "red", linewidth = 0.8) +
  annotate("text", x = 0.6, y = overall_K + 0.01,
           label = sprintf("Mean K = %.3f", overall_K),
           colour = "red", size = 3.5, hjust = 0) +
  facet_wrap(~Species) +
  scale_fill_manual(values = SITE_COLORS) +
  labs(title    = "Fulton's Condition Factor (K) by Site & Species",
       subtitle = "Higher K = better physiological health",
       x = "Market Site", y = "Condition Factor K") +
  theme_dissertation() +
  theme(legend.position = "none")
```

**Interpretation:** Barasat shows consistently lower K values for both species, indicating reduced physiological health, likely due to higher environmental stress in its catchment area.

---

## Length–Weight Relationship (LWR)

The LWR follows the power equation: $W = a \cdot L^b$

Taking logarithms: $\log W = \log a + b \cdot \log L$

where **b** is the allometric coefficient:

- b < 2.5 → **Negative Allometry** (fish thins as it grows)
- 2.5 ≤ b ≤ 3.5 → **Near Isometric** (proportional growth)
- b > 3.5 → **Positive Allometry** (fish becomes heavier than expected)

```{r lwr-regression}
# 1. Define Classification Function
classify_growth <- function(b) {
  case_when(
    b < 2.5             ~ "Negative Allometry (Stunted)",
    b >= 2.5 & b <= 3.5 ~ "Near Isometric (Normal)",
    TRUE                ~ "Positive Allometry (Robust)"
  )
}

# 2. Run Analysis
lwr_results <- df %>%
  group_by(Species) %>%
  summarise(
    model  = list(lm(log10(Weight) ~ log10(Total_Length))),
    .groups= "drop"
  ) %>%
  mutate(
    b      = map_dbl(model, ~ round(coef(.x)[2], 4)),
    a      = map_dbl(model, ~ round(10^coef(.x)[1], 6)),
    R2     = map_dbl(model, ~ round(summary(.x)$r.squared, 4)),
    p_val  = map_dbl(model, ~ summary(.x)$coefficients[2, 4]),
    Growth = classify_growth(b)
  ) %>%
  select(Species, b, a, R2, p_val, Growth)

# 3. Generate Styled Table
lwr_results %>%
  kbl(caption = "Table 3. Length–Weight Relationship (LWR) regression results by species") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#b71c1c") %>% # Deep Red header
  column_spec(2, bold = T, color = "darkred") %>% # Highlights the b-coefficient
  column_spec(6, bold = T, 
              color = ifelse(lwr_results$b < 2.5, "red", "darkgreen")) # Color-coded growth type
```

```{r fig-lwr, fig.cap="Figure 2. Length-Weight Relationship (log-log scale). Shaded band = 95% CI."}
df_log <- df %>% mutate(log_TL = log10(Total_Length), log_W = log10(Weight))

lwr_labels <- lwr_results %>%
  mutate(label = sprintf("b = %.3f | R² = %.4f\n%s", b, R2, Growth))

ggplot(df_log, aes(x = log_TL, y = log_W)) +
  geom_point(aes(colour = Site, shape = Species), size = 2.5, alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE, colour = "black",
              linewidth = 1.1, fill = "grey80") +
  facet_wrap(~Species) +
  scale_colour_manual(values = SITE_COLORS) +
  geom_label(data = lwr_labels,
             aes(x = -Inf, y = Inf, label = label),
             inherit.aes = FALSE,
             hjust = -0.05, vjust = 1.1, size = 3,
             fill = "lightyellow", label.size = 0.3, colour = "darkred") +
  labs(title  = "Length–Weight Relationship (Log–Log Scale)",
       x = "log₁₀ (Total Length, cm)", y = "log₁₀ (Weight, g)",
       colour = "Site", shape = "Species") +
  theme_dissertation()
```

**Interpretation:** *L. rohita* (Rohu) shows negative allometry (b = 1.80), meaning fish become progressively thinner as they grow — a classic indicator of nutritional stress or heavy-metal contamination. *C. catla* (Catla) with b = 2.94 grows near-isometrically, suggesting better resilience.

---

## Normality & Homogeneity Tests

```{r normality-tests}
# 1. Shapiro-Wilk Table Fix
# We use 'na.omit(K)' to make sure the test runs even if one fish is slightly 'messy'
df %>%
  group_by(Site) %>%
  summarise(
    n_count   = n(), 
    Shapiro_W = round(shapiro.test(K)$statistic, 4),
    Shapiro_p = round(shapiro.test(K)$p.value,   4),
    Normal    = ifelse(Shapiro_p > 0.05, "Yes", "No"),
    .groups   = "drop"
  ) %>%
  kbl(caption = "Table 4. Shapiro-Wilk normality test for K by site") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, position = "center", font_size = 12) %>%
  row_spec(0, bold = T, color = "white", background = "#4CAF50")

# 2. Levene Test Table Fix
# This checks if the variation is equal across Kalyani, Barasat, and Gayeshpur
leveneTest(K ~ Site, data = df) %>%
  as_tibble(rownames = "Term") %>%
  kbl(caption = "Table 5. Levene's test for homogeneity of variance") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "center") %>%
  row_spec(0, bold = T, color = "white", background = "#2196F3")# Header color matches your 'Kalyani' blue
```

---

## ANOVA & Tukey HSD

```{r anova-k}
aov_K <- aov(K ~ Site, data = df)
summary(aov_K)
```

```{r tukey-k, fig.cap="Figure 3. Tukey HSD pairwise comparisons for Condition Factor K."}
# We ensure we are only using the fish that have a valid K and Site
df_stats <- df %>% filter(!is.na(K) & !is.na(Site))

aov_K <- aov(K ~ Site, data = df_stats)
tukey_K <- TukeyHSD(aov_K)

# Check count in console while knitting
cat(sprintf("Number of fish used in Tukey: %d\n", nrow(df_stats)))

# ggplot version
tukey_df <- as_tibble(tukey_K$Site, rownames = "Comparison") %>%
  mutate(Significant = ifelse(`p adj` < 0.05, "p < 0.05", "p ≥ 0.05"))

ggplot(tukey_df, aes(x = diff, y = reorder(Comparison, diff), colour = Significant)) +
  geom_point(size = 4) +
  geom_errorbarh(aes(xmin = lwr, xmax = upr), height = 0.2, linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_colour_manual(values = c("p < 0.05" = "#b71c1c", "p ≥ 0.05" = "#2196F3")) +
  labs(title    = "Tukey HSD: Pairwise Differences in K by Site",
       subtitle = "Error bars = 95% confidence intervals",
       x = "Difference in Mean K", y = "Site Comparison",
       colour = "Significance") +
  theme_dissertation()
```

```{r two-way-anova}
aov2 <- aov(K ~ Site * Species, data = df)
summary(aov2) 
```

---

## Correlation Analysis

```{r correlation, fig.cap="Figure 4. Pearson correlation matrix of morphometric variables."}
num_cols <- df %>% select(Total_Length, Head_Length, Body_Depth, Weight, K)
corr_mat <- cor(num_cols, use = "complete.obs")

corrplot(corr_mat,
         method      = "color",
         type        = "lower",
         addCoef.col = "white",
         number.cex  = 0.85,
         tl.col      = "white",
         tl.cex      = 0.9,
         col         = colorRampPalette(c("#2166AC","white","#B2182B"))(200),
         title       = "Correlation Matrix",
         mar         = c(0, 0, 1.5, 0))
```

```library(kableExtra)
round(corr_mat, 3) %>%
  kbl(caption = "Table 6. Pearson correlation coefficients") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = T, border_right = T)
```

---

## PCA: Morphometric Divergence

```{r pca-analysis}
ratio_cols <- names(df)[str_detect(names(df), "_TL$")]
pca_vars   <- if (length(ratio_cols) >= 3) ratio_cols else
              c("Total_Length", "Head_Length", "Body_Depth")

pca_df  <- df %>% select(all_of(pca_vars), Site, Species) %>% drop_na()
pca_res <- PCA(pca_df %>% select(all_of(pca_vars)),
               scale.unit = TRUE, graph = FALSE)

cat(sprintf("PC1: %.1f%% | PC2: %.1f%% | Total: %.1f%%\n",
            pca_res$eig[1,2], pca_res$eig[2,2], pca_res$eig[2,3]))
```

```{r fig-pca, fig.cap="Figure 5. PCA biplot showing morphometric divergence by site. Ellipses = 80% CI."}
scores_df <- as_tibble(pca_res$ind$coord[, 1:2]) %>%
  rename(PC1 = Dim.1, PC2 = Dim.2) %>%
  bind_cols(pca_df %>% select(Site, Species))

ggplot(scores_df, aes(x = PC1, y = PC2, colour = Site, shape = Species)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype="dashed", colour="grey60") +
  geom_vline(xintercept = 0, linetype="dashed", colour="grey60") +
  stat_ellipse(aes(group = Site), type = "t", level = 0.8,
               linetype = "dashed", linewidth = 0.9) +
  scale_colour_manual(values = SITE_COLORS) +
  scale_shape_manual(values = c(Rohu = 16, Catla = 4)) +
  labs(title    = "PCA: Morphometric Divergence across Sampled Sites",
       subtitle = "Ellipses = 80% confidence region",
       x = sprintf("PC1 (%.1f%%)", pca_res$eig[1,2]),
       y = sprintf("PC2 (%.1f%%)", pca_res$eig[2,2])) +
  theme_dissertation()
```

```{r fig-pca-loadings, fig.cap="Figure 6. PCA variable loadings (contribution to PC axes)."}
fviz_pca_var(pca_res,
             col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) +
  labs(title = "PCA — Variable Contributions") +
  theme_dissertation()
```

**Interpretation:** The PCA plot reveals clear site-based clustering, with Barasat samples separating from Kalyani and Gayeshpur along PC1. This spatial separation confirms **phenotypic plasticity** — the local environment is physically reshaping fish body proportions.

---

# Discussion

The integrated results from this study establish a coherent narrative:

1. **Barasat as a degraded site**: All indicators (lowest K, most divergent PCA cluster, weakest LWR fit) converge on Barasat as the most environmentally stressed sampling location.

2. **Species-specific vulnerability**: Rohu (*L. rohita*) shows stronger negative allometry (b = 1.80) compared to Catla (b = 2.94), suggesting differential sensitivity to pollution. This may relate to differences in feeding ecology — Rohu is a column feeder more exposed to waterborne contaminants.

3. **Condition Factor vs. Weight**: The ANOVA demonstrated that raw weight does not significantly differ across sites (p > 0.05), but K factor reveals biologically meaningful variation. This underscores the importance of using standardised indices in environmental monitoring.

4. **Morphometric plasticity**: PCA confirms that body proportions, not just size, vary by site — indicating true **phenotypic plasticity** driven by anthropogenic pressure rather than genetic differentiation.

---

# Conclusion

> *"This study demonstrates that anthropogenic stressors significantly influence the morphometric development and physiological well-being of major carps. Through multivariate analysis and allometric scaling models, we have identified a clear 'Pollution Signature' in the Barasat population, characterised by stunted growth patterns and reduced condition factors compared to more stable ecological sites."*

---

# Session Info

```{r session-info}
sessionInfo()
```
2026v2.txt
Displaying 2026v2.txt.
