---
---
Title: "Comparative Morphometric & Physiological Analysis of *Labeo rohita* and *Catla catla* across West Bengal Market Sites"
Author: "ANANYA SAMADDER | UID No: 22BSZ70039 | HONOURS IN ZOOLOGY | DEPARTMENT OF BIOSCIENCES"
date: "`r format(Sys.Date(13th of September; 2026), '%B %Y')`"


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  dpi     = 150
)
```

---

# Abstract

This study investigates morphometric and physiological variation between two commercially important Indian carp species — *Labeo rohita* (Rohu) and *Catla catla* (Catla) — sampled from three market sites in West Bengal: Kalyani, Barasat, and Gayeshpur. Using Fulton's Condition Factor (K), Length–Weight Relationship (LWR) regressions, Principal Component Analysis (PCA), Pearson correlation, and ANOVA with Tukey HSD post-hoc tests, we find that condition factor differed significantly by species but not by site, with L. rohita showing negative allometric growth and C. catla growing closer to isometrically. PCA showed some site-associated clustering in body proportions, independent of the non-significant site effect on condition factor. 

---

# Introduction

Major carps such as *Labeo rohita* and *Catla catla* are ecologically and economically vital to West Bengal's aquaculture and fisheries sector. Morphometric variation across spatially separated populations can reveal signatures of regional stock heterogeneity and phenotypic variation. Comparing market sites offers a practical basis for examining condition and growth patterns across locations. Market-sourced specimens offer a convenient way to sample across multiple nearby sites, though this reflects the sampled catch at each market rather than a direct survey of the originating water bodies, and sites here were selected for logistical accessibility rather than a priori environmental hypotheses.

**Research Objectives:**

1. Quantify physiological health (K) across three market sites.
2. Determine growth patterns (allometry) using Length–Weight Relationships.
3. Explore multivariate morphometric structure across sites using PCA.
4. Test for statistical differences in condition factor across species and sites using ANOVA and Tukey HSD, and examine associations among morphometric variables using Pearson correlation.

---

# Materials & Methods

## Data Collection

Fish specimens were collected from three market sites:

| Site | Location | n (Catla) | n (Rohu) |
|------|----------|-----------|----------|
| Kalyani | Nadia District | ~12 | ~12 |
| Barasat | North 24 Parganas | ~10 | ~12 |
| Gayeshpur | Nadia District | ~12 | ~12 |

Measurements recorded for each specimen: Total Length (TL, cm), Head Length (HL, cm), Body Depth (BD, cm), Dorsal Fin Length (DFL), Pectoral Fin Length (PFL), Anal Fin Length (AFL), and Weight (g).

## Statistical Methods

| Method | Purpose |
|--------|---------|
| Fulton's K | Condition factor (physiological index) |
| LWR (log–log regression) | Growth allometry (slope b) |
| Pearson Correlation | Association between morphometric variables |
| One-Way ANOVA | Test for site effect on K |
| Tukey HSD | Post-hoc pairwise comparisons |
| Two-Way ANOVA | Test for site, species, and site × species effects on K |
| PCA | Exploration of multivariate morphometric structure |
| Post-hoc power analysis | Assess statistical power of the site-effect test given sample size |

---

# Setup & Data Loading

```{r libraries}
library(tidyverse)
library(ggpubr)
library(car)
library(multcomp)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(rstatix)
library(knitr)

SITE_COLORS    <- c(Kalyani = "#2196F3", Barasat = "#FF9800", Gayeshpur = "#4CAF50")
SPECIES_COLORS <- c(Rohu = "#1a237e", Catla = "#b71c1c")
SITES          <- c("Kalyani", "Barasat", "Gayeshpur")

theme_dissertation <- function() {
  theme_classic(base_size = 12) +
    theme(
      plot.title    = element_text(face = "bold", size = 13, hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5, colour = "grey40"),
      axis.title    = element_text(face = "bold"),
      legend.title  = element_text(face = "bold"),
      strip.background = element_rect(fill = "grey92", colour = NA),
      strip.text    = element_text(face = "bold")
    )
}
```

```{r load-data}
catla <- read_csv("final_catla.csv") %>% mutate(Species = "Catla")
rohu  <- read_csv("final_rohu.csv")  %>% mutate(Species = "Rohu")

df <- bind_rows(catla, rohu) %>%
  rename(Total_Length = TL, Head_Length = HL, Body_Depth = BD) %>%
  mutate(
    Site    = factor(Site,    levels = SITES),
    Species = factor(Species, levels = c("Rohu", "Catla"))
  )

cat(sprintf("Total records: %d\n", nrow(df)))
```

## Descriptive Statistics

```{r descriptive_statistics}
library(kableExtra)
df %>%
  group_by(Species, Site) %>%
  summarise(
    n           = n(),
    TL_mean     = round(mean(Total_Length, na.rm = TRUE), 2),
    TL_sd       = round(sd(Total_Length,   na.rm = TRUE), 2),
    Weight_mean = round(mean(Weight, na.rm = TRUE), 2),
    Weight_sd   = round(sd(Weight,   na.rm = TRUE), 2),
    .groups     = "drop"
  ) %>%
  kbl(caption = "Table 1. Descriptive statistics by Species and Site") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#455A64") %>% # Professional Slate Grey header
  column_spec(1, bold = T, border_right = T) %>% # Bolds Species and adds a separator
  column_spec(2, italic = T) # Italics the Site names for a professional look
```

---

# Results

## Fulton's Condition Factor (K)

Fulton's K was calculated as:

$$K = \frac{W}{L^3} \times 100$$

where W = weight (g) and L = total length (cm).

```{r condition-factor}

# 1. Calculate K
df <- df %>%
  mutate(K = (Weight / Total_Length^3) * 100)

# 2. Generate Styled Table
df %>%
  group_by(Species, Site) %>%
  summarise(
    K_mean = round(mean(K, na.rm=TRUE), 4),
    K_sd   = round(sd(K,   na.rm=TRUE), 4),
    .groups = "drop"
  ) %>%
  kbl(caption = "Table 2. Mean Condition Factor (K) by Species and Site") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#607D8B") %>% # Professional Blue-Grey header
  column_spec(3, bold = T, color = "darkblue") # Highlights the Mean K values for quick reading
```

```{r fig-condition-factor, fig.cap="Figure 1. Fulton's Condition Factor across sites. Dashed line = overall mean K."}
overall_K <- mean(df$K, na.rm = TRUE)

ggplot(df, aes(x = Site, y = K, fill = Site)) +
  geom_boxplot(outlier.shape = 21, alpha = 0.75, width = 0.6, colour = "grey30") +
  geom_hline(yintercept = overall_K, linetype = "dashed",
             colour = "red", linewidth = 0.8) +
  annotate("text", x = 0.6, y = overall_K + 0.01,
           label = sprintf("Mean K = %.3f", overall_K),
           colour = "red", size = 3.5, hjust = 0) +
  facet_wrap(~Species) +
  scale_fill_manual(values = SITE_COLORS) +
  labs(title    = "Fulton's Condition Factor (K) by Site & Species",
       subtitle = "Higher K = better physiological health",
       x = "Market Site", y = "Condition Factor K") +
  theme_dissertation() +
  theme(legend.position = "none")
```
**Interpretation:** Barasat shows a descriptively lower mean K for Rohu and a descriptively higher mean K for Catla relative to the other sites, but the one-way ANOVA found no significant site effect (p = 0.788) and no pairwise Tukey comparison reached significance. This pattern should not be described as a confirmed site-level difference.

---

---

## Length–Weight Relationship (LWR)

The LWR follows the power equation: $W = a \cdot L^b$

Taking logarithms: $\log W = \log a + b \cdot \log L$

where **b** is the allometric coefficient:

- b < 2.5 → **Negative Allometry** (fish thins as it grows)
- 2.5 ≤ b ≤ 3.5 → **Near Isometric** (proportional growth)
- b > 3.5 → **Positive Allometry** (fish becomes heavier than expected)

```{r lwr-regression}
# 1. Define Classification Function
classify_growth <- function(b) {
  case_when(
    b < 2.5             ~ "Negative Allometry (Stunted)",
    b >= 2.5 & b <= 3.5 ~ "Near Isometric (Normal)",
    TRUE                ~ "Positive Allometry (Robust)"
  )
}

# 2. Run Analysis
lwr_results <- df %>%
  group_by(Species) %>%
  summarise(
    model  = list(lm(log10(Weight) ~ log10(Total_Length))),
    .groups= "drop"
  ) %>%
  mutate(
    b      = map_dbl(model, ~ round(coef(.x)[2], 4)),
    a      = map_dbl(model, ~ round(10^coef(.x)[1], 6)),
    R2     = map_dbl(model, ~ round(summary(.x)$r.squared, 4)),
    p_val  = map_dbl(model, ~ summary(.x)$coefficients[2, 4]),
    Growth = classify_growth(b)
  ) %>%
  select(Species, b, a, R2, p_val, Growth)

# 3. Generate Styled Table
lwr_results %>%
  kbl(caption = "Table 3. Length–Weight Relationship (LWR) regression results by species") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  ) %>%
  row_spec(0, bold = T, color = "white", background = "#b71c1c") %>% # Deep Red header
  column_spec(2, bold = T, color = "darkred") %>% # Highlights the b-coefficient
  column_spec(6, bold = T, 
              color = ifelse(lwr_results$b < 2.5, "red", "darkgreen")) # Color-coded growth type
```

```{r fig-lwr, fig.cap="Figure 2. Length-Weight Relationship (log-log scale). Shaded band = 95% CI."}
df_log <- df %>% mutate(log_TL = log10(Total_Length), log_W = log10(Weight))

lwr_labels <- lwr_results %>%
  mutate(label = sprintf("b = %.3f | R² = %.4f\n%s", b, R2, Growth))

ggplot(df_log, aes(x = log_TL, y = log_W)) +
  geom_point(aes(colour = Site, shape = Species), size = 2.5, alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE, colour = "black",
              linewidth = 1.1, fill = "grey80") +
  facet_wrap(~Species) +
  scale_colour_manual(values = SITE_COLORS) +
  geom_label(data = lwr_labels,
             aes(x = -Inf, y = Inf, label = label),
             inherit.aes = FALSE,
             hjust = -0.05, vjust = 1.1, size = 3,
             fill = "lightyellow", label.size = 0.3, colour = "darkred") +
  labs(title  = "Length–Weight Relationship (Log–Log Scale)",
       x = "log₁₀ (Total Length, cm)", y = "log₁₀ (Weight, g)",
       colour = "Site", shape = "Species") +
  theme_dissertation()
```

**Interpretation:** *Labeo rohita* (Rohu) shows negative allometry (b = 1.9450, R² = 0.4537),indicating that body mass accumulates at a slower rate relative to linear length extension. while *Catla catla* (Catla) grows closer to isometrically (b = 2.6853, R² = 0.6421). showing more proportional mass-to-length scaling.These distinct scaling coefficients reflect the inherent morphological growth patterns of each species within the sampled size range.
---


---

## Normality & Homogeneity Tests

```{r normality-tests}
# 1. Shapiro-Wilk Table Fix
df %>%
  group_by(Site) %>%
  summarise(
    Shapiro_W = round(shapiro.test(K)$statistic, 4),
    Shapiro_p = round(shapiro.test(K)$p.value,   4),
    Normal    = ifelse(shapiro.test(K)$p.value > 0.05, "Yes", "No"),
    .groups   = "drop"
  ) %>%
  kbl(caption = "Table 4. Shapiro-Wilk normality test for K by site") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, position = "center", font_size = 12) %>%
  row_spec(0, bold = T, color = "white", background = "#4CAF50") # Header color matches your 'Gayeshpur' green

# 2. Levene Test Table Fix
leveneTest(K ~ Site, data = df) %>%
  as_tibble(rownames = "Term") %>%
  kbl(caption = "Table 5. Levene's test for homogeneity of variance") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "center") %>%
  row_spec(0, bold = T, color = "white", background = "#2196F3") # Header color matches your 'Kalyani' blue
```

---

## ANOVA & Tukey HSD

```{r anova-k}
aov_K <- aov(K ~ Site, data = df)
summary(aov_K)
```

```{r tukey-k, fig.cap="Figure 3. Tukey HSD pairwise comparisons for Condition Factor K."}
tukey_K <- TukeyHSD(aov_K)
print(tukey_K)

# ggplot version of Tukey
tukey_df <- as_tibble(tukey_K$Site, rownames = "Comparison") %>%
  mutate(Significant = ifelse(`p adj` < 0.05, "p < 0.05", "p ≥ 0.05"))

ggplot(tukey_df, aes(x = diff, y = reorder(Comparison, diff), colour = Significant)) +
  geom_point(size = 4) +
  geom_errorbarh(aes(xmin = lwr, xmax = upr), height = 0.2, linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_colour_manual(values = c("p < 0.05" = "#b71c1c", "p ≥ 0.05" = "#2196F3")) +
  labs(title    = "Tukey HSD: Pairwise Differences in K by Site",
       subtitle = "Error bars = 95% confidence intervals",
       x = "Difference in Mean K", y = "Site Comparison",
       colour = "Significance") +
  theme_dissertation()
```

```{r two-way-anova}
aov2 <- aov(K ~ Site * Species, data = df)
summary(aov2) 
```

```{r power-analysis}
library(pwr)

SS_site <- 0.0136
SS_resid <- 1.9118
eta_sq <- SS_site / (SS_site + SS_resid)
cohen_f <- sqrt(eta_sq / (1 - eta_sq))

pwr.anova.test(k = 3, n = 12, f = cohen_f, sig.level = 0.05)
```


---

## Correlation Analysis

```{r correlation, fig.cap="Figure 4. Pearson correlation matrix of morphometric variables."}
num_cols <- df %>% select(Total_Length, Head_Length, Body_Depth, Weight, K)
corr_mat <- cor(num_cols, use = "complete.obs")

corrplot(corr_mat,
         method      = "color",
         type        = "lower",
         addCoef.col = "white",
         number.cex  = 0.85,
         tl.col      = "white",
         tl.cex      = 0.9,
         col         = colorRampPalette(c("#2166AC","white","#B2182B"))(200),
         title       = "Correlation Matrix",
         mar         = c(0, 0, 1.5, 0))
```

```library(kableExtra)
round(corr_mat, 3) %>%
  kbl(caption = "Table 6. Pearson correlation coefficients") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = T, border_right = T)
```

---

## PCA: Morphometric Divergence

```{r pca-analysis}
ratio_cols <- names(df)[str_detect(names(df), "_TL$")]
pca_vars   <- if (length(ratio_cols) >= 3) ratio_cols else
              c("Total_Length", "Head_Length", "Body_Depth")

pca_df  <- df %>% select(all_of(pca_vars), Site, Species) %>% drop_na()
pca_res <- PCA(pca_df %>% select(all_of(pca_vars)),
               scale.unit = TRUE, graph = FALSE)

cat(sprintf("PC1: %.1f%% | PC2: %.1f%% | Total: %.1f%%\n",
            pca_res$eig[1,2], pca_res$eig[2,2], pca_res$eig[2,3]))
```

```{r fig-pca, fig.cap="Figure 5. PCA biplot showing morphometric divergence by site. Ellipses = 80% CI."}
scores_df <- as_tibble(pca_res$ind$coord[, 1:2]) %>%
  rename(PC1 = Dim.1, PC2 = Dim.2) %>%
  bind_cols(pca_df %>% select(Site, Species))

ggplot(scores_df, aes(x = PC1, y = PC2, colour = Site, shape = Species)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype="dashed", colour="grey60") +
  geom_vline(xintercept = 0, linetype="dashed", colour="grey60") +
  stat_ellipse(aes(group = Site), type = "t", level = 0.8,
               linetype = "dashed", linewidth = 0.9) +
  scale_colour_manual(values = SITE_COLORS) +
  scale_shape_manual(values = c(Rohu = 16, Catla = 4)) +
  labs(title    = "PCA: Morphometric Divergence across Sampled Sites",
       subtitle = "Ellipses = 80% confidence region",
       x = sprintf("PC1 (%.1f%%)", pca_res$eig[1,2]),
       y = sprintf("PC2 (%.1f%%)", pca_res$eig[2,2])) +
  theme_dissertation()
```

```{r fig-pca-loadings, fig.cap="Figure 6. PCA variable loadings (contribution to PC axes)."}
fviz_pca_var(pca_res,
             col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) +
  labs(title = "PCA — Variable Contributions") +
  theme_dissertation()
```

**Interpretation:**  Statistical Validation: Shapiro-Wilk and Levene's tests confirm normal distribution and variance homogeneity for most sites, validating the use of parametric tests. ANOVA Results: Species had a significant effect on condition factor (p = 0.000454); site (p = 0.758) and the site × species interaction (p = 0.859) did not. Tukey HSD found no significant pairwise site differences. Correlation: body weight and total length were strongly correlated (r ≈ 0.74–0.75). PCA: PC1 and PC2 together explained 61.8% of variance, with some site-associated clustering along PC1 driven by body-depth and head-length loadings - this pattern did not correspond to a statistically significant site effect.
Furthermore, a post-hoc power analysis (Cohen's $f = 0.084$, group sample sizes $n = 10\text{–}12$) indicated only ~6.8% power to detect the observed effect size at $\alpha = 0.05$, indicating no significant site-level differences were found, and subtle geographic variations cannot be entirely ruled out due to the limited sample size.

---

# Discussion

The integrated results from this study establish a narrative:

1. Primary influence of species identity: The two-way ANOVA revealed a significant species effect on $K$ ($p = 0.000454$), whereas neither site ($p = 0.758$) nor the site $\times$ species interaction ($p = 0.859$) showed significance, indicating that species identity is the principal source of variance.

2. Species-specific allometry: Rohu displays negative allometric growth ($b = 1.9450$), whereas Catla demonstrates near-isometric development ($b = 2.6853$), representing an inherent species-level differences.

3. Condition factor sensitivity: Fulton’s $K$ provides a standardized metric that accounts for length-weight relationships, allowing for consistent physiological condition assessment across individuals.

4. Principal component patterns: Although PCA indicated minor site-associated separation along PC1 based on body-depth and head-length loadings, this trend did not align with a statistically significant site effect on $K$, pointing toward future work with expanded sample sizes rather than confirmed population divergence.

---

# Conclusion

> "This study documents a statistically significant species-level difference in condition factor and growth allometry between *Labeo rohita* and *Catla catla* sampled from West Bengal market sites. Site-level differences in condition factor were not statistically significant in this dataset, though PCA indicated some site-associated clustering in body proportions warranting further investigation in future studies."

---

# Session Info

```{r session-info}
sessionInfo()
```
